%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% title:
% authors:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\fullver{0}
\def\submission{1}
\input{defs}


\addbibresource{notes.bib}


\title{\textbf{Designing Secure Cryptography}}

\author{
  Cornell CS 6831
}


\ifnum\submission=1
\date{}
\fi



\begin{document}

\maketitle


\section*{Notes}

\tnote{A couple notes for myself, will remove later.}
\begin{itemize}
  \item Fix a pseudocode language, including randomness, and an associated model
    of computation. Treat memory usage as well as run times of algorithms. We
    will most often treat memory as subservient to run-time, bounding the former
    by the latter. But special treatments are needed.
    The goal is to have a simplified abstraction, but which
    provides good predictions of running code.  Probably look at ``Careful with
    Composition'' paper as starting point.

 \item Algorithms are simply pseudocode that take input and produce an output.
    Define what it means for algorithm to be runnable. Define adversaries
    as algorithms and require they be runnable. Comment on how this interacts
    with (and rules out) non-uniform reductions, other such things.

  \item  Algorithms can be parameterized, but this is just notational sugar. The result
    must still be runnable. They may have oracles, and this defines some
    interface.  Give conventions on resource usage of algorithms: number of
    oracle queries, run time (with oracle queries unit cost), etc. Discuss how
    runnable algorithms can then be composed.

  \item Associate probability space to any algorithm, be pedantic here since we
    may want to prove some basic stuff via direct manipulation of probability
    space (i.e., coin-counting arguments).


  \item Introduce security definitions as special algorithms called games. The
    game is just another algorithm parameterized by some adversary, scheme, etc.
    Thus typically what we define as a game is actually a family of algorithms,
    one for each instantiation of things. (How much notion of a template do we
    need? Is this confusing?) It is used to measure adversarial advantage, a
    which is a measure of success for an adversary.

  \item Discuss the viewpoint underlying all the above. Security models are
    coarse abstractions of real cryptographic systems. They must be coarse for
    us to do certain kinds of analysis. This means that what we prove in our
    formalizations do not typically apply to real systems. But they are a good
    heuristic: schemes for which we have good analyses tend to provide better
    security. We will therefore judge the value of models by their utility in
    helping us build cryptographic schemes that resist attackers in practice.
    We will include examples and discussions reinforcing this viewpoint along
    the way. (One example to treat right away is asymptotics. Could be good
    heuristic, but is often quite poor, and gives little advice about how to set security
    parameters meaningfully.)

  \item Discuss assumptions. These are formalized as games. What separates them
    from security definitions is contextual. They are ``lower level'' and
    less related to the goals of cryptographic protocols.

  \item Reductions are runnable algorithms.

  \item Using reductions to set security parameters. Maybe interleave some of
    this discussion with first couple of chapters
\end{itemize}


\newpage
\input{intro}
\newpage
\input{notation}
%\newpage
%\input{ciphers}
\newpage
\input{ciphers-draft}
\newpage
\input{prf}
\input{prp/all}
\newpage
\input{cryptanalysis}
\newpage
\input{freqanalysis}
\newpage
\input{tweakciphers}
\newpage
\input{symenc}
\newpage
\input{authenc}
\newpage
\input{msgauth}
\newpage
\input{genericcomp}
\newpage
\input{advanced-ae}
\newpage
\input{aepractice}
\newpage
\input{hashfunctions}
\newpage
\input{hashfunctions2}
\newpage
\input{pke}

\printbibliography

%\bibliographystyle{plain}
%\bibliography{notes}


\end{document}
